---
title: "restData patch"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

In *Week 3* of the [Getting and Cleaning Data](https://www.coursera.org/learn/data-cleaning) Coursera course, the lectures *Summarizing Data* and *Creating New Variables* use a [link to a CSV](https://data.baltimorecity.gov/api/views/k5ry-ef3g/rows.csv?accessType=DOWNLOAD) file that no longer exists. Luckily, the link to the HTML data demonstrated on another slide was accessed and cached.

Following this patch, it is possible recreate `restData` from this HTML.

```{r}
library(dplyr)
library(rvest)

# read restaurants.html into restData
cache <- "https://github.com/oeryania/datasciencecoursera/raw/main/restaurants.html"
restData <- cache %>% read_html %>% html_table %>% as.data.frame %>% mutate(.keep = "none",
               name = as.factor(NAME),
            zipCode = as.integer(sub("(.*)-$", "-\\1", ZIPCODE)),
       neighborhood = as.factor(NGHBRHD),
    councilDistrict = as.integer(CNCLDST),
     policeDistrict = as.factor(PLCDST),
         Location.1 = as.factor(paste(ADDRESS, "\n", CITY, ", ", STATE, "\n", sep = ""))
    )
```

`restData` is now ready for further exploration.

---

<br><br><br><br>

## Data Exploration
### Summarizing Data

#### Slide 4. Look at a bit of the data

```{r}
head(restData,n=3)
```

```{r}
tail(restData,n=3)
```

#### Slide 5. Make summary

```{r}
summary(restData)
```

#### Slide 6. More in depth information

```{r}
str(restData)
```

#### Slide 7. Quantiles of quantitative variables

```{r}
quantile(restData$councilDistrict,na.rm=TRUE)
```

```{r}
quantile(restData$councilDistrict,probs=c(0.5,0.75,0.9))
```

#### Slide 8. Make table

```{r}
table(restData$zipCode,useNA="ifany")
```

#### Slide 9. Make table

```{r}
table(restData$councilDistrict,restData$zipCode)
```

#### Slide 10. Check for missing values

```{r}
sum(is.na(restData$councilDistrict))
```

```{r}
any(is.na(restData$councilDistrict))
```

```{r}
all(restData$zipCode > 0)
```

#### Slide 11. Row and column sums

```{r}
colSums(is.na(restData))
```

```{r}
all(colSums(is.na(restData))==0)
```

#### Slide 12. Value with specific characteristics

```{r}
table(restData$zipCode %in% c("21212"))
```

```{r}
table(restData$zipCode %in% c("21212","21213"))
```

#### Slide 13. Value with specific characteristics

```{r}
restData[restData$zipCode %in% c("21212","21213"),]
```

### Creating New Variables

#### Slide 6. Subsetting variables

```{r}
restData$nearMe = restData$neighborhood %in% c("Roland Park", "Homeland")
table(restData$nearMe)
```

#### Slide 7. Creating binary variables

```{r}
restData$zipWrong = ifelse(restData$zipCode < 0, TRUE, FALSE)
table(restData$zipWrong,restData$zipCode < 0)
```

#### Slide 8. Creating categorical variables
```{r}
restData$zipGroups = cut(restData$zipCode,breaks=quantile(restData$zipCode))
table(restData$zipGroups)
```

```{r}
table(restData$zipGroups,restData$zipCode)
```

#### Slide 9. Easier cutting
```{r}
library(Hmisc)
restData$zipGroups = cut2(restData$zipCode,g=4)
table(restData$zipGroups)
```

#### Slide 10. Creating factor variables
```{r}
restData$zcf <- factor(restData$zipCode)
restData$zcf[1:10]
```

```{r}
class(restData$zcf)
```

#### Slide 13. Using the mutate function
```{r}
restData2 = mutate(restData,zipGroups=cut2(zipCode,g=4))
table(restData$zipGroups)
```
